Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-10-10 paper/comoving-followup.tex
sky -> tan
update table
comments to myself

\newcommand{\npairsobs}{311}
\newcommand{\ncomoving}{127}
\newcommand{\nraveoverlap}{154}
\newcommand{\nravecomoving}{59}
\newcommand{\ncomovingtotal}{186}

@@

\begin{abstract}
% Context
Members of widely-separated stellar binaries ($\gtrsim 0.1$~pc) or loosely bound
stellar associations and clusters are born with small spreads in kinematics and
chemistry, but disrupt and dissociate on timescales comparable to their orbital
time within the Milky Way.
The kinematic properties of comoving star pairs---bound wide binaries or unbound
but coeval stars---are therefore sensitive to both the smooth tidal field and
gravitational perturbation spectrum of the Milky Way.
% Aims
% Thus far, a few hundred confirmed comoving pairs with separations as large as
% $8~\pc$ have been discovered in the solar neighborhood using
% \project{Hipparcos} astrometry combined with radial velocity surveys.
In previous work, we identified $>4,000$ candidate comoving pairs using only
astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog and
find a large number of pairs with separations $\gtrsim 1~\pc$.
Without radial velocity measurements for the vast majority of these pairs, the
false-positive rate at these separations is significant ($\sim 50\%$).
% Methods
Here we present results from our own low-resolution radial-velocity survey of a
random sample of the candidate pairs to identify and validate true comoving star
pairs.
% Results
Of the \npairsobs\ observed comoving pairs, we confirm \ncomoving\ comoving
pairs with separations as large as $\approx 10~\pc$, our original search limit.
Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
the number of high-probability pairs per separation decreases with increasing
separation (number per log separation is uniform).
From $\approx 0.5~\pc$ to $10~\pc$, the number per separation is approximately
uniform (number per log separation increases).
% Conclusions
We confirm the discovery of a population of comoving star pairs at very large
separations, suggesting that disrupted wide binaries and stellar associations
remain approximately spatially coherent.
\end{abstract}

% The confirmed discovery of this population of coeval, comoving stars will
% enable future dynamical tests of dark matter and XX of star formation XX.
% as predicted from simulations of the disruption and evolution of wide binary
% stars. Gaia DR2 -> how many do we expect? Can we say anything? Out to what
% distance and separation for a typical MS star can we determine they are bound
% assuming high-res spectra to get very precise RVs?

@@

To verify the existence of a population of very wide-separation comoving pairs,
we first need to identify a large sample of high-probability comoving stars.
This will already be possible for the brightest stars with \gaia\ \DR{2}, but
obtaining even larger samples will require significant efforts to obtain radial
velocities for candidate comoving pairs.
In this work, we present a catalog of comoving star pairs discovered through
low-resolution spectroscopic follow-up of candidate comoving stars selected from
inferences using astrometric measurements from the \tgas\ catalog alone
(\citealt{Oh:2017}).

We describe the data and observations in \sectionname~\ref{sec:data}.
Details of the spectroscopic reduction and velocity measurements are described
in the Appendix.
We present our main results in \sectionname~\ref{sec:results} and
\sectionname~\ref{sec:discussion}.

@@

\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=0.8\linewidth]{sample_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
    blue histogram) and all comoving star pair members observed in this
    follow-up effort (black points); colored lines connect candidate comoving
    star pairs.
    \tmass\ (\citealt{Skrutskie:2006}) $J$-band magnitudes are obtained
    from the Gaia science archive.
    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
    \citealt{Oh:2017}).
    \label{fig:sample-cmd}}
\end{figure}

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
We performed a model selection between two hypotheses: that the velocities of
the two stars in a pair are identical (case 1) or that they are indepndent (case
2).
We marginalized over the (unknown) true velocity and distance of each star to
obtain the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$.
The likelihoods appropriately take into account the reported uncertainties and
covariances of the astrometric measurements in the \tgas\ catalog
as well as projection effect in comparing spherical
velocities for star pairs separated by large angles on the sky.

After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$,
we compute the marginalized likelihood ratio for all stars within $10~\pc$ and
with a difference in tangential velocity $\Delta v_\perp < 10~\kms$ for each
star in the \tgas\ catalog.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars, $\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
However, we have also found a large number ($\approx$4,000) of
mutually-exclusive linked pairs that increase in number with increasing
log-separation from $\sim 1$~\pc.
For a full description of the selection method, see \citealt{Oh:2017}.

To select candidate pairs for follow-up spectroscopic observations, we only
consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
distance.
With the SNR cut used above and given the limitations of the \tgas\ catalog
itself, \tgas\ is $\approx$90\% complete to $200~\pc$ (\citealt{Bovy:2017}).
From this source target list, we randomly observed targets with airmass $\sec z
< 1.5$ during our observing run.
\figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
observed in this observing run (markers) plotted over the density of all \tgas\
stars within $200~\pc$.
Candidate comoving pairs are connected by (colored) lines.

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
For calibration sources, we use radial velocity values reported in the source
catalogs (\citealt{Bensby:2014,Pinsonneault:2014}).
For all other sources, following previous work (\citealt{Shaya:2011}), we search
for previous RV measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  (\citealt{Ginsburg:2016}) to retrieve records from the database, which do not
  contain uncertainties but do contain a quality flag. We only retain RVs that
  have a quality flag set to `A' and adopt a $5~\kms$ uncertainty for all such
  values.
\end{enumerate}
% See: paper/figures/Compare with previous observations.ipynb
We find that 155 of the observed stars (mainly calibration targets) have
reliable prior measurements from the literature search described above.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution (right panel), we compute the median absolute deviation
(MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
adopt as our global systematic error for all absolute radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision (typically $\approx 5$--$10~\kms$) that
matters.

@@

Our model is expressed in terms of the magnitude of the true (latent) 3D
velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
and the (global) fraction of pairs that are comoving, $f$.
For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
f)$, we use a mixture model with two components
\begin{eqnarray}
    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
\end{eqnarray}
where $p_{\rm co}$, the probability distribution function (pdf) of true comoving
pairs, and $p_{\rm field}$, the pdf of randomly paired field stars.
Using a mixture model and inferring the mixture weight is a trick to avoid
explicitly having to sample over the probabilities for each individual object
(candidate comoving pair): this expression is equivalent to the marginal
likelihood under a model in which every candidate comoving pair has an
additional latent parameter (a binary flag, $q_n$) that specifies which mixture
component it belongs to (\citealt{Hogg:2010,Foreman-Mackey:2014}).
We later use this fact to compute the probability that each pair is truly
comoving, or, more precisely, belongs to the mixture component $p_{\rm co}$.

The form for both mixture components ($p_{\rm co}$ and $p_{\rm field}$) is taken
to be the probability distribution of the magnitude of the 3D difference between
two vectors under the assumption that the two vectors were drawn independently
from an isotropic, 3D Gaussian with dispersion $\sigma$.
For the three-dimensional velocity vectors we consider here, the distribution is
Maxwellian in form.
That is,
\begin{eqnarray}
    f(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right]
        \label{eq:vel-diff-dist}\\
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{v} \given \bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& \mathcal{N}(\psi \given 0, \sigma^2_{\rm co})\\
    p_{\rm field}(\psi) &=& f(\psi \,;\, \sigma_{\rm field})
\end{eqnarray}
where $\sigma_{\rm co}$ and $\sigma_{\rm field}$ are defined below,
$\mathcal{N}(\bs{v} \given \bs{m}, \mat{\Sigma})$ represents a normal
distribution over $\bs{v}$ with mean vector $\bs{m}$ and covariance matrix
$\mat{\Sigma}$, $\mathcal{N}(x \given m, \sigma^2)\,[a,b]$ is a truncated normal
distribution with mean $m$ and variance $\sigma^2$ defined on the domain
$[a,b]$, and $\mat{I}_3$ is the 3 by 3 identity matrix.

@@

The posterior probability over the fraction of true comoving star pairs, $f$, is
then given by
\begin{equation}
    p(f \given \{\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2}\}_N)
        \propto p(f) \, \prod_n^N
        p(\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2} \given f)
\end{equation}
where $(\bs{\eta}_{n1}, \varpi_{n1})$ and $(\bs{\eta}_{n2}, \varpi_{n2})$ refer
to the first and second star in the $n$-th comoving pair, respectively, and  $N$
is the total number of candidate comoving star pairs.

@@

At the smallest tangential separations of our candidate pairs, ($s_{\rm tan}
\approx 10^{-3}~\pc$), if the two stars are bound, the expected velocity
difference is a few $\kms$ (depending on mass).
We use the phrase ``tangential separation'' (and symbol $s_{\rm tan}$) to refer
to the chord length separation between the two pairs assuming the two stars are
at the same distance (the smaller distance of the two, $d_{\rm min}$):
\begin{equation}
    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
\end{equation}
where $\theta$ is the angular extent of the pair.
We set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function at zero)
so that binary stars with very small but finite velocity differences are
considered comoving.
In principle, this ``tolerance'' on finite velocity difference should decrease
with increasing separation, at least out to the radius at which most wide
binaries are disrupted.
At the unbinding radius, the instantaneous velocity difference between the two
stars should be comparable to the orbital velocity of the binary at unbinding
(typically $\approx$0.1--$0.5~\kms$; \citealt{Jiang:2010}), but the difference
may also increase steadily through dynamical heating.
Rather than constructing a model for the full process in this work, we use a
constant velocity tolerance ($\sigma_{\rm co}$) and defer to future work a full
population analysis that also infers the separation distribution of comoving
stars, which could instead express the model as a phase-space distribution
function so that the velocity difference tolerance effectively becomes a
function of separation.
We note that even with this tolerance, it is very improbable that two stars
sampled at random from the field population will have a 3D velocity difference
$<1~\kms$.

@@

\figurename~\ref{fig:mixture-model-demo} shows the probability distributions for
each mixture model component, $p_{\rm field}(\psi)$ and $p_{\rm co}(\psi)$,
normalized so that the peak of the distribution is at 1.
As noted above, even though the field dispersion is set to $\sigma_{\rm field}
\approx 18~\kms$, most randomly paired stars will have 3D velocity differences
(i.e. $\psi$) significantly larger than the value of the dispersion.

To generate samples in $f$, we use Markov Chain Monte Carlo sampling as
implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
We use 28 walkers and keep track of the value of the marginal likelihood under
each mixture model component for each sample at every iteration step.
We run the sampler for 256 burn-in steps from a small ball of initial positions
centered on $f=0.5$, then restart the walkers from the final position and run
for an additional 1024 steps.
With the samples in $f$ and the mixture model component likelihoods, we compute
the per-pair probabilities of belonging to the comoving component (see the final
section of \citealt{Foreman-Mackey:2014}).

@@

We use the model described in \sectionname~\ref{sec:pair-probs} to constrain the
fraction of stars consistent with being comoving given 3D velocities for
candidate comoving pairs (a) with radial velocities obtained and described in
this work---\npairsobs\ pairs, and (b) with radial velocities previously
measured and released in \DR{5} of the Radial Velocity Experiment
(\acronym{RAVE}; \citealt{Kunder:2017})---202 pairs.
Using the posterior samples in $f$, we compute the probability (under this
model) that each given pair is comoving including the radial velocity
measurements.
\figurename~\ref{fig:f-samples} shows histograms of posterior pdf samples in $f$
for pairs with our radial velocity measurements (blue, darker histogram) and
with \acronym{RAVE} radial velocities (orange, lighter histogram); in either
case, $f \approx 40\%$ implying a contamination rate $1-f \approx 60\%$.

@@

\figurename~\ref{fig:separation} shows histograms of the tangential separation
for all pairs with radial velocities from this work and the \acronym{RAVE}
survey, with each pair weighted by the probability that it is comoving.
Interestingly, the separation distribution of the probable comoving pairs
appears to be uniform with increasing separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

@@

To see whether the separation distribution depends on heliocentric distance, we
also plot the tangential separation of all comoving pairs with comoving
probability $> 50\%$ as a function of heliocentric distance in
\figurename~\ref{fig:separation-scatter}.
The top histogram again shows the tangential separation distribution for the
joint sample (now only for pairs with $> 50\%$ comoving probability).
The scatter plot shows tangential separation and the pair distances, and
the right histogram shows the distance distribution of these high-probability
pairs.
Reassuringly, there do not appear to be any trends of the separation
distribution with distance.

@@

\figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
diagram for pairs with radial velocities measured in this work with comoving
probability $> 0.5$.
As would be expected if all of the pairs are coeval, the majority of pairs lie
along isochrones in color-magnitude space, with the exception of a few visible
outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs in the next section
(\sectionname~\ref{sec:interesting-pairs}).

@@

From their surface gravities and temperatures, the component stars of this pair
appear to be an A-type dwarf (HD 92278) and a slightly less massive G-type
subgiant (HD 95088).
This pair is separated by $4.5^\circ$ on the sky (corresponding to $\approx
8.6~\pc$ in tangential separation, see \tablename~\ref{tbl:highlighted-pairs})
and have consistent distances within their uncertainties of $\approx 110 \pm
5~\pc$.
At first glance, it seems unlikely that the two stars could be coeval, as the
more massive A star has not begun its post-main-sequence evolution.
One possible explanation is that this pair is a false-positive in three
dimensions, in which case they are unassociated but just appear to be comoving
with small relative velocity at present day.

>>>>

\newcommand{\npairsobs}{311}
\newcommand{\nhighprob}{124}
\newcommand{\nraveoverlap}{154}
\newcommand{\nravecomoving}{59}
\newcommand{\ncomovingtotal}{186}

@@

\begin{abstract}
% Context
Members of widely-separated stellar binaries ($\gtrsim 0.1$~pc) or loosely bound
stellar associations and clusters are born with small spreads in kinematics and
chemistry, but are thought to physically dissociate over short timescales.
The kinematic properties of comoving stars---bound, wide or unbound but coeval
systems---are therefore sensitive to both the smooth tidal field and the
spectrum of massive perturbers (e.g., stellar remnants, dark matter, giant
molecular clouds) in the Milky Way.
% Aims
% Thus far, a few hundred confirmed comoving pairs with separations as large as
% $8~\pc$ have been discovered in the solar neighborhood using
% \project{Hipparcos} astrometry combined with radial velocity surveys.
In previous work, we identified $>4,000$ candidate comoving star pairs using
only astrometric data from the Tycho-Gaia Astrometric Solution (TGAS) catalog
and found a large number of candidate comoving pairs with separations $\gtrsim
1~\pc$.
Without radial velocity measurements for the vast majority of these pairs, the
false-positive rate at large separations is significant ($\sim 60\%$).
% Methods
Here we present results from our own low-resolution radial-velocity survey of a
random sample of the candidate pairs to identify high-probability comoving star
pairs.
% Results
Of the \npairsobs\ observed comoving pairs, \nhighprob\ comoving pairs have
probabilities $50\%$ with separations as large as $\approx 10~\pc$, our original
search limit.
Between separations of $10^{-3}~\pc$ to $\approx 0.5~\pc$,
the number of high-probability pairs per separation decreases with increasing
separation.
From $\approx 0.5~\pc$ to $10~\pc$, the number per separation is approximately
uniform.
% Conclusions
We thus confirm the discovery of a population of comoving star pairs at very
large separations, suggesting that disrupted wide binaries and stellar
associations remain approximately spatially coherent.
\todo{something about stellar models and chemical tagging}
\end{abstract}

@@

Our primary goal in this work is to identify a sample of high-probability, very
wide-separation comoving pairs to be followed up with high-resolution
spectroscopy in future work to determine detailed chemical abundances.
This will be possible for the brightest stars with \gaia\ \DR{2}, but obtaining
even larger samples will require significant efforts to obtain radial velocities
for candidate comoving pairs at fainter magnitudes.
Here we present a catalog of comoving star pairs discovered through our own
low-resolution spectroscopic follow-up of candidate comoving stars selected from
inferences using astrometric measurements from the \tgas\ catalog alone
(\citealt{Oh:2017}).
For each pair, we provide a probability of being comoving as determined by
modeling the 3D velocity information for each pair, with radial velocities
either from this work, or from \DR{5} of the Radial Velocity Experiment
(\acronym{RAVE}; \citealt{Kunder:2017}).

We describe the data and observations in \sectionname~\ref{sec:data}, then
present a probabilistic model to infer the comoving probability for each pair
with radial velocity measurements in \sectionname~\ref{sec:pair-probs}.
We present our main results in \sectionname~\ref{sec:results} and
\sectionname~\ref{sec:discussion}.
Details of the spectroscopic reduction and radial velocity measurements for our
observations are described in the Appendix.

@@

\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=0.8\linewidth]{sample_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram for all 195,294 \tgas\ sources with \acronym{2MASS}
    photometry within $200~\pc$ (background blue histogram) and all comoving
    star pair members observed in this follow-up effort (black points); colored
    lines connect candidate comoving star pairs.
    \tmass\ (\citealt{Skrutskie:2006}) $J$-band magnitudes are obtained
    from the Gaia science archive.
    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
    \citealt{Oh:2017}).
    \label{fig:sample-cmd}}
\end{figure}

@@

In previous work, we have identified a sample of high-confidence, candidate
comoving star pairs (\citealt{Oh:2017}) using only astrometric information from
the \tgas\ catalog (\citealt{Michalik:2015,Gaia-Collaboration:2016a}).
These comoving pairs were selected using a cut on a marginalized likelihood
ratio computed for all nearly co-spatial pairs of stars.
We performed a model selection between two hypotheses: that the velocities of
the two stars in a pair are identical (case 1) or that they are indepndent (case
2).
We marginalized over the (unknown) true 3D velocity and distance of each star to
obtain the ratio of fully marginalized likelihoods,
$\mathcal{L}_1/\mathcal{L}_2$.
The likelihoods appropriately take into account the reported uncertainties and
covariances of the astrometric measurements in the \tgas\ catalog as well as
projection effects with comparing spherical velocities for star pairs separated
by large angles on the sky.

After a parallax signal-to-noise ratio (SNR) cut, $\varpi/\sigma_\varpi > 8$,
for each star, we compute the marginalized likelihood ratio for all nearby stars
within $10~\pc$ and with a difference in tangential velocity $\Delta v_\perp <
10~\kms$.
After applying a conservative cut on the likelihood ratio motivated by computing
the same likelihood ratio for random pairings of stars, $\ln
\mathcal{L}_1/\mathcal{L}_2 > 6$, we found 13,058 comoving star pairs with
physical separations between $10^3~\au$ and $10~\pc$.
Of these pairs, there are 10,606 unique sources, implying that a significant
fraction of the identified pairs are members of larger groups.
However, we have also found a large number ($\approx$4,000) of
mutually-exclusive linked pairs that increase in number with increasing
log-separation from $\sim 1$~\pc.
For a full description of the selection method, see \citealt{Oh:2017}.

To select candidate pairs for follow-up spectroscopic observations, we only
consider mutually exclusive star pairs within $200~\pc$ in Heliocentric
distance.
With the SNR cut used above and given the limitations of the \tgas\ catalog
itself, \tgas\ is $\approx$90\% complete to $200~\pc$ (\citealt{Bovy:2017}).
From this source target list, we randomly observed targets with airmass $\sec z
< 1.5$ during our observing run.
\figurename~\ref{fig:sample-cmd} shows a color-magnitude diagram of all targets
observed in this observing run (markers) plotted over the density of 195,294
\tgas\ stars within $200~\pc$ with \acronym{2MASS} photometry.
Candidate comoving pairs are connected by (arbitrarily colored) lines.

@@

Several of our targets and all of our calibration targets have previously
measured radial velocities (RVs).
We can therefore estimate the absolute velocity uncertainties by comparing our
derived radial velocities with high-quality prior measurements.
For calibration sources, we use radial velocity values reported in the source
catalogs (\citealt{Bensby:2014,Pinsonneault:2014}).
For all other sources, following previous work (\citealt{Shaya:2011}), we search
for previous RV measurements (in decreasing order of priority) with:
\begin{enumerate}
  \item The Geneva-Copenhagen Survey of the Solar Neighborhood
  (\citealt{Nordstrom:2004}).
  \item The Catalogue of Radial Velocities of Galactic Stars with Astrometric
  Data, the Second Version (\citealt{Kharchenko:2007}).
  \item The General Catalogue of Mean Radial Velocities
  (\citealt{Barbier-Brossat:2000}).
  \item The SIMBAD database (\citealt{Wenger:2000}). We use \package{astroquery}
  (\citealt{Ginsburg:2016}) to retrieve records from the database, which do not
  contain uncertainties but do contain a quality flag. We only retain RVs that
  have a quality flag set to `A' and adopt a $5~\kms$ uncertainty for all such
  values.
\end{enumerate}
% See: paper/figures/Compare with previous observations.ipynb
We find that 155 of the observed stars (mainly calibration targets) have
reliable prior measurements from the literature search described above.
\figurename~\ref{fig:compare-previous} shows the distribution of velocity
differences between our measurements and literature values.
From this distribution (right panel), we compute the median absolute deviation
(MAD) and estimate the standard deviation to be $\approx 20~\kms$, which we
adopt as our global systematic error for all \emph{absolute} radial velocities.
We note, however, that for computing \emph{relative} velocities for pairs of
nearby stars, it is the RV precision (typically $\approx 5$--$10~\kms$) that
matters.

@@

Our model is expressed in terms of the magnitude of the true (latent) 3D
velocity difference for each pair, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$,
and the (global) fraction of pairs that are comoving, $f$.
For the distribution of true 3D velocity difference magnitudes, $p(\psi \given
f)$, we use a mixture model with two components
\begin{eqnarray}
    p(\psi \given f) &=& f\,p_{\rm co}(\psi) + (1-f)\,p_{\rm field}(\psi)
    \label{eq:mixture}
\end{eqnarray}
where $p_{\rm co}$ is the probability distribution function (pdf) of true
comoving pairs, and $p_{\rm field}$ is the pdf of randomly paired field stars.
Using a mixture model and inferring the mixture weight is a trick to avoid
explicitly having to sample over the probabilities for each individual object
(candidate comoving pair): this expression (\eqname~\ref{eq:mixture}) is
equivalent to the marginal likelihood under a model in which every candidate
comoving pair has an additional latent parameter (a binary flag, $q_n$) that
specifies which mixture component it belongs to
(\citealt{Hogg:2010,Foreman-Mackey:2014}).
We later use this fact to compute the probability that each pair is truly
comoving, or, more precisely, belongs to the mixture component $p_{\rm co}$.

For the field mixture component, $p_{\rm field}$, we use the probability
distribution of the magnitude of the 3D difference between two vectors under the
assumption that the two vectors were drawn independently from an isotropic, 3D
Gaussian with dispersion $\sigma_{\rm field}$.
For the three-dimensional velocity vectors we consider here, the distribution is
Maxwellian in form.
For the comoving mixture component, $p_{\rm co}$, we use a truncated normal
distribution centered on 0 with dispersion $\sigma_{\rm co}$ defined over the
domain $\psi \in [0, \infty]$.
To summarize,
\begin{eqnarray}
    \bs{v_1}, \bs{v_2} &\sim& \mathcal{N}(\bs{v} \given \bs{0}, \sigma^2\,\mat{I}_3) \\
    p_{\rm co}(\psi) &=& \mathcal{N}(\psi \given 0, \sigma^2_{\rm co})\,[0, \infty]\\
    p_{\rm field}(\psi) &=& g(\psi \,;\, \sigma_{\rm field})\\g(\psi \,;\, \sigma) &=& \left[2\sqrt {\pi}\, \sigma^3 \right]^{-1} \,
        \psi^2 \, \ \exp\left[-\frac {\psi^2} {4\, \sigma^2} \right]
        \label{eq:vel-diff-dist}
\end{eqnarray}
where $\mathcal{N}(\bs{v} \given \bs{m}, \mat{\Sigma})$ represents a normal
distribution over $\bs{v}$ with mean vector $\bs{m}$ and covariance matrix
$\mat{\Sigma}$, $\mathcal{N}(x \given m, \sigma^2)\,[a,b]$ is a truncated normal
distribution with mean $m$ and variance $\sigma^2$ defined on the domain
$[a,b]$, $\mat{I}_3$ is the 3 by 3 identity matrix, and the values of
$\sigma_{\rm co}$ and $\sigma_{\rm field}$ are set below,

@@

The posterior probability over the fraction of true comoving star pairs, $f$, is
then given by
\begin{equation}
    p(f \given \{\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2}\}_N)
        \propto p(f) \, \prod_n^N
        p(\bs{\eta}_{n1}, \bs{\eta}_{n2}, \varpi_{n1}, \varpi_{n2} \given f)
\end{equation}
where $(\bs{\eta}_{n1}, \varpi_{n1})$ and $(\bs{\eta}_{n2}, \varpi_{n2})$ refer
to the velocity and parallax data for the first and second star in the $n$-th
comoving pair, respectively, and $N$ is the total number of candidate comoving
star pairs.

@@

At the smallest tangential separations of our candidate pairs, ($s_{\rm tan}
\approx 10^{-3}~\pc$), if the two stars are bound, the expected velocity
difference is a few $\kms$ (depending on mass).
We use the phrase ``tangential separation'' (and symbol $s_{\rm tan}$) to refer
to the chord length separation between the two pairs assuming the two stars are
at the same distance (the smaller distance of the two, $d_{\rm min}$):
\begin{equation}
    s_{\rm tan} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
\end{equation}
where $\theta$ is the angular extent of the pair.
We therefore set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function
at zero) so that binary stars with very small but finite velocity differences
are considered comoving.
In principle, this ``tolerance'' on finite velocity difference should decrease
with increasing separation, at least out to the radius at which most wide
binaries are disrupted.
At the unbinding radius, the instantaneous velocity difference between the two
stars should be comparable to the orbital velocity of the binary at unbinding
(typically $\approx$0.1--$0.5~\kms$; \citealt{Jiang:2010}), but the difference
may also increase steadily through dynamical heating.
Rather than constructing a model for this full process in this work, we use a
constant velocity tolerance ($\sigma_{\rm co}$) and defer to future work a full
population analysis that also infers the separation distribution of comoving
stars, which could instead express the model as a phase-space distribution
function so that the velocity difference tolerance effectively becomes a
function of separation.
We note that even with this tolerance, it is very improbable that two stars
sampled at random from the field population will have a 3D velocity difference
$<1~\kms$.

@@

\figurename~\ref{fig:mixture-model-demo} shows the probability distributions for
each mixture model component, $p_{\rm field}(\psi)$ and $p_{\rm co}(\psi)$,
normalized so that the peak of each distribution is at 1.
As noted above, even though the field dispersion is set to $\sigma_{\rm field}
\approx 18~\kms$, most randomly paired stars will have 3D velocity differences
(i.e. $\psi$) significantly larger than the value of the dispersion.

To generate posterior samples in $f$, we use Markov Chain Monte Carlo sampling
as implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
We use 28 walkers and keep track of the value of the marginal likelihood under
each mixture model component for each sample at every iteration step.
We run the sampler for 256 burn-in steps from a small ball of initial positions
centered on $f=0.5$, then restart the walkers from the final position and run
for an additional 1024 steps.
With the samples in $f$ and the mixture model component likelihoods, we compute
the per-pair probabilities of belonging to the comoving component (see the final
section of \citealt{Foreman-Mackey:2014}).

@@

We use the model described in \sectionname~\ref{sec:pair-probs} to constrain the
fraction of stars consistent with being comoving given 3D velocities for the
candidate comoving pairs (a) with radial velocities obtained and described in
this work---\npairsobs\ pairs, and (b) with radial velocities previously
measured and released in \DR{5} of the \acronym{RAVE} survey
(\citealt{Kunder:2017})---202 pairs.
Using the posterior samples in $f$, we compute the probability (under this
model) that each given pair is comoving including the radial velocity
measurements.
\figurename~\ref{fig:f-samples} shows histograms of posterior pdf samples in $f$
for pairs with our radial velocity measurements (blue, darker histogram) and
with \acronym{RAVE} radial velocities (orange, lighter histogram); in either
case, $f \approx 40\%$ implying a contamination rate $1-f \approx 60\%$.

@@

\figurename~\ref{fig:separation} shows histograms of the tangential separation
for all pairs with radial velocities from this work and the \acronym{RAVE}
survey, with each pair weighted by the probability that it is comoving.
Interestingly, the separation distribution appears to be uniform with increasing
separation above $\approx1~\pc$.
This suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).
However, high resolution spectroscopy is needed to determine precise stellar
parameters, ages, and high-precision radial velocities for all of our
high-probability comoving pairs to further validate the sample.

@@

To see whether the separation distribution depends on heliocentric distance,
which could indicate a problem with our statistical inferences, we also plot the
tangential separation of all comoving pairs with comoving probability $> 50\%$
as a function of heliocentric distance in
\figurename~\ref{fig:separation-scatter}.
The top histogram again shows the tangential separation distribution for the
joint sample (now only for pairs with $> 50\%$ comoving probability).
The scatter plot shows tangential separation and the pair distances, and
the right histogram shows the distance distribution of these high-probability
pairs.
Reassuringly, there do not appear to be any trends of the separation
distribution with distance.

@@

\figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
diagram for high-probability ($>50\%$) pairs with radial velocities measured in
this work.
As would be expected if all of the pairs are coeval, the majority of pairs lie
along isochrones in color-magnitude space, with the exception of a few visible
outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs in the next section
(\sectionname~\ref{sec:interesting-pairs}).

@@

From their surface gravities and temperatures, the component stars of this pair
appear to be an A-type dwarf (HD 92278) and a slightly less massive G-type
subgiant (HD 95088).
This pair is separated by $4.5^\circ$ on the sky (corresponding to $\approx
8.6~\pc$ in tangential separation, see \tablename~\ref{tbl:highlighted-pairs})
and have consistent distances within their uncertainties of $\approx 110 \pm
5~\pc$.
At first glance, it seems impossible that the two stars could be coeval, given
that the more massive A star has not begun its post-main-sequence evolution.
One possible explanation is that this pair is a false-positive in three
dimensions, in which case they are unassociated but just appear to be comoving
with small relative velocity at present day.
